### 2014-11-02 fundamentals/impl.tex

A key innovation in \Cyclus is implementing this generic \gls{API} and modular 
architecture into a suite of dynamically loadable ``plugin'' libraries.  Though 
common in modern software architecture, such a ``plugin'' pattern has 
previously not been implemented in nuclear fuel cycle simulators in the 
literature.  In \Cyclus, this implementation maximizes the potential of the clean, 
object-oriented \gls{API} and modern build system already described.

Dynamically-loadable libraries are the primary mechanism for extending \Cyclus' capability. 
This approach provides encapsulation. The core of the code operates
completely independently from the individual plugin libraries. Thus, any
customization or extension is implemented only in the loadable
library.

@@

In large simulations there may be many material objects changing frequently.
Material decay can also contribute significantly to such changes.  In order to
help avoid unnecessary runtime performance and database space impacts,
compositions in \Cyclus have some special features, such as immutability.
Immutable compositions allow multiple material objects to all hold
references to the same composition object safely.  When operations are
performed on a material that do not change the composition, any new resulting
materials hold a reference to the same, originating composition object.
Although the decay operation generates a new, separate composition (preserving
immutability), a reference to the new composition is placed in a cache shared
by the new originating compositions. The next time the originating
composition is decayed, the cache will be checked for a prior calculation
with the same cumulative decay time.  If one exists, the cached composition is
used.  This cache is shared by all compositions that transitively come from
the same origin.  Composition immutability in concert with decay history
caching help eliminate many redundant decay calculations in addition to
reducing the total number of composition objects.  Since each composition
object is only recorded in the database once, significant space savings also
occurs. \TODO{do we want a diagram helping show this composition caching?}

>>>>

A key innovation in \Cyclus is implementing this generic \gls{API} and modular 
architecture into a suite of dynamically loadable plug-in libraries.  Though 
common in modern software architecture, such a plug-in pattern has 
previously not been implemented in nuclear fuel cycle simulators in the 
literature.  In \Cyclus, this implementation maximizes the potential of the clean, 
object-oriented \gls{API} and modern build system already described.

Dynamically loadable libraries are the primary mechanism for extending \Cyclus' capability. 
This approach provides encapsulation: the core of the code operates
completely independently from the individual plug-in libraries. Thus, any
customization or extension is implemented only in the loadable
library.

@@

In large simulations there may be many material objects changing frequently.
Material decay can also contribute significantly to such changes.  In order to
help avoid unnecessary runtime performance and database space impacts,
compositions in \Cyclus have some special features, such as immutability.
Immutable compositions allow multiple material objects to all hold
references to the same composition object safely.  When operations are
performed on a material that do not alter the composition, any new resulting
materials hold a reference to the same, original composition object.
Although the decay operation generates a new, separate composition (preserving
immutability), a reference to the new composition is placed in a cache shared
by the new and original compositions. The next time the original
composition is decayed, this cache will be checked for a prior calculation
with the same cumulative decay time.  If one exists, the cached composition is
used.  This cache is shared by all compositions that transitively come from
the same origin.  Composition immutability in concert with decay history
caching help eliminate many redundant decay calculations in addition to
reducing the total number of composition objects.  Since each composition
object is only recorded in the database once, significant space savings also
occurs. \TODO{do we want a diagram helping show this composition caching?}

### 2014-11-02 fundamentals/intro.tex

The \Cyclus Nuclear Fuel Cycle Simulator framework and 
its modeling ecosystem answer this need for a next generation nuclear fuel cycle 
simulator. Dynamic simulation of nuclear fuel cycles with discrete agents 
and per-nuclide species resolution support 
levels of detail appropriate for a varied community of analysts
interested in a range of analysis types. Additionally, \Cyclus employs an 
open development process that supports user access and 
encourages developer extensions.

>>>>

The \Cyclus Nuclear Fuel Cycle Simulator framework and 
its modeling ecosystem answer this need for a next generation nuclear fuel cycle 
simulator. This dynamic simulator employs discrete agents, time, and materials 
up to isotopic resolution in order to support a range of analysis types, levels of 
detail, and fuel cycle analysts.
Additionally, \Cyclus employs a modular architecture and 
open development process that supports user access and 
encourages developer extensions.

### 2014-11-02 fundamentals/qa.tex

Validation on the other hand may be defined as the question, 
``Is \Cyclus the correct tool?''
Since \Cyclus is alone in its class as an agent-based fuel cycle simulator, longitudinal 
validation is not possible. Still code-to-code comparisons with fuel cycle
simulators with other modeling paradigms are underway, if nascent. However, such 
exercises are more likely to bring into relief the differences between the modeling
paradigms than be useful for \gls{QA} and validation.

Lastly, \gls{UQ} is a process that is used on specific simulation
instantiations to statistically determine its quality. Since the \Cyclus 
kernel requires an input file written by the user, \gls{UQ} is a process that applies 
to the running of many \Cyclus simulations. The kernel itself, as a generic fuel cycle 
simulator, cannot make reliable statements about the uncertainty of a particular fuel 
cycle. Any \gls{UQ} results are applicable only to the scenario that is under 
analysis. \Cyclus is not a panacea for due dillegence on behalf of the fuel 
cycle modeler. Thus, \gls{UQ} is envisioned to be beyond the scope of core development 
and in the domain of the expert user.

@@

It is important to note that before a proposed
code change is allowed into \Cyclus, all of the tests must pass
and the change must be covered by a test, either new or existing. This is 
matter of policy for the developers. Additionally, test code is equally likely 
as the code it is testing to contain errors. To prevent an infinite recursion 
of testing-the-tests, it is assumed that diligence and one level of testing 
is sufficient to meet the aims of \gls{QA}.

@@

Integration tests are any tests that combine multiple elements of the 
\Cyclus interface and test that they work correctly with each other.  By analogy, 
simply because the gears (units) are made correctly does not imply that the 
clock (integration) will run smoothly, run at all, or give the correct time.
In \Cyclus and \Cycamore, integration tests are performed by running sample
simulations and verifying that results are what would be expected ahead of 
time. Pre-canned input files have thus been constructed. Their results 
are inspected and compared via nose \cite{pellerin_nose_2007}, a Python test framework.
Python will run cyclus as a subprocess for each integration test. In this
way, \Cyclus code units are tested in the full context that they will be
executed. This category of testing is especially useful for ensuring that 
major \Cyclus components are functioning as expected.

@@

Regression tests are tests that ensure significant changes do not 
occur over the course of \Cyclus development. Such a change is called a 
\emph{regression} because the new version is almost always wrong.
Regression tests are implemented similarly to integration tests.
Nose is used to execute fully fledged \Cyclus simulations whose results
are then compared. In this category, however, the comparison is done against 
the output of the same input file when run with a previous version of \Cyclus, 
typically the last released version.
In some sense, regression tests are `dumb' in that they do 
not care about the contents of a simulation being correct, only whether or not 
it changed. They thus pick up glaring time-sensitive mistakes that may 
not appear elsewhere though they do not purport to make claims about the
quality of the physics modeled. The \Cycamore project is the primary location of
regression tests. This is because the archetypes here are sophisticated enough
to make interesting input files. Only a handful ($<10$) of time steps are needed 
for regression testing.

@@

All of the public interface (the \gls{API}) must be documented as part of \Cyclus
\gls{QA} policy. This certifies that the intention for a code unit is communicated 
in way that is separate from its implementation. This policy is important 
because implementations are often wrong; software implementations are unreliable 
resources about themselves. Truly trustworthy communication about what 
software should be doing comes from the author explicitly stating the 
goals of an \gls{API} in prose form. Future development of the code is then able to 
compare the implementation versus the documentation to discern whether or not they 
are in agreement. Documentation is thus a secondary higher-level information 
stream.  In \Cyclus, this information is aggregated together into static 
websites with the Doxygen \cite{van_heesch_doxygen:_2008} and Sphinx 
\cite{brandl_sphinx_2014} 
tools. These websites are placed online at \url{http://fuelcycle.org}.

@@

Version control is the history-preserving mechanism by which the provenance of 
a code is recorded. \Cyclus uses a well-established tool called git 
\cite{software_freedom_conservancy_git_2014}
for this purpose. Git is currently the most popular utility in the recent wave
of \emph{distributed version control}. Briefly stated, this means that every 
user has a complete local copy of the \emph{repository} (the version control
term for the collection of all possible histories).
Though there and many praises and criticisms of git, one 
feature that the tool performs exceptionally well is the notion of code branches.
These are distinct pathways in the history that branch out of a mainline source
tree and then may be \emph{merged} back into the mainline. Multiple simultaneous
branches may exist at all points in time. Every change to the code is recorded
in the history, along with metadata such as the author, a timestamp, and an 
accompanying message. Thus 
it is possible for \Cyclus to accurately replay the entirety of who did what to the
code when.

\Cyclus uses a strategy known as \emph{git flow} 
\cite{kalliamvakou_code-centric_2014} 
to manage topical branches where individuals develop code, the mainline development
branch (simply known as develop), and the stable branch (called master).
In addition to these rules and delineation, \Cyclus also imposes rules on 
when it is acceptable from a quality control standpoint to merge code from 
the personal topical branches into develop. All software at this stage 
must have sufficient and passing tests and comprehensive documentation. Without
these core pieces, proposed changes are not allowed into \Cyclus. A schematic of 
the development stages between reporting a bug and merging the fix is shown in 
Figure \ref{fig:gitprocess}.

\begin{figure}[htbp]
\begin{center}
\includegraphics[height=0.6\textheight]{./images/gitprocess}
\end{center}
\caption{A version-control enabled process for finding, reporting, and fixing a bug 
can take place within an issue tracker alongside multiple stages of code 
testing and review. In the \Cyclus case, the final review stage includes unit, 
regression, and integration testing as well as manual code inspection.}
\label{fig:gitprocess}
\end{figure}

@@

In any multi-person software project, there is a tension between how individuals
wish to write code. Every person tends to have their own custom style. To remedy this
coding style guides are the software analogy to those for written language,
such as the Chicago Manual of Style. \Cyclus strictly enforces the use of the 
Google C++ Style Guide \cite{weinberger_google_2008} for all software contributions.
This means that all developers of \Cyclus nominally write \Cyclus code in the same 
way.  This homogenization may be a hurdle to new developers but ultimately 
improves code legibility. Adoption of a single style guide also settles
timeless arguments over what style is `the best.' Having only one means that 
there is a common enemy for all developers to dislike equally.

@@

\emph{\acrfull{CI}} the idea that software should be tested and validated 
as it is being developed, rather than as a final stage in a longer development 
cycle.  Under \gls{CI}, every pull request is tested independently immediately after 
it is proposed. Such tests should be run on all officially supported platforms. 
\Cyclus uses a \gls{CI} platform called Polyphemus \cite{scopatz_polyphemus_2014}.

@@

\gls{CI} has become a necessary but not sufficient addition to the \Cyclus \gls{QA}
system. Its merit is that it prevents bad code from being implemeted. However, \Cyclus will always
require human eyes and hands to let good code in.

>>>>

Validation on the other hand may be defined as the question, 
``Is \Cyclus the correct tool?''
Since \Cyclus is alone in its class as an agent-based fuel cycle simulator, longitudinal 
validation is not possible. Still code-to-code comparisons with fuel cycle
simulators with other modeling paradigms are underway, if nascent. However, such 
exercises are more likely to bring into relief the differences between the modeling
paradigms than be useful for QA and validation.

Lastly, uncertainty quantification is a process that is used on specific simulation
instantiations to statistically determine its quality. Since the \Cyclus 
kernel requires an input file written by the user, UQ is a process that applies 
to the running of many \Cyclus simulations. The kernel itself, as a generic fuel cycle 
simulator, cannot make reliable statements about the uncertainty of a particular fuel 
cycle. Any UQ results are applicable only to the scenario that is under 
analysis. \Cyclus is not a panacea for due dillegence on behalf of the fuel 
cycle modeler. Thus, UQ is envisioned to be beyond the scope of core development 
and in the domain of the expert user.

@@

It is important to note that before a proposed
code changes is allowed into main stream \Cyclus, all of the tests must pass
and the change must be covered by a test, either new or existing. This is 
matter of policy for the developers. Additionally, test code is equally likely 
as the code it is testing to contain errors. To prevent an infinite recursion 
of testing-the-tests, it is assumed that diligence and one level of testing 
is sufficient to meet the aims of the quality assurance.

@@

Integration tests are any tests that combine multiple elements of the 
\Cyclus interface and test that they work correctly with each other.  By analogy, 
simply because the gears (units) are made correctly does not imply that the 
clock (integration) will run smoothly, run at all, or give the correct time.
In \Cyclus and \Cycamore, integration tests are performed by running sample
simulations and verifying that results are what would be expected ahead of 
time. Pre-canned input files have thus been constructed. Their results 
are inspected and compared via nose \cite{pellerin_nose_2007}, a Python test framework.
Python will run cyclus as a subprocess for each integration test. In this
way cyclus code units are tested in the full context that they will be
executed. This category of testing is especially useful for ensuring that 
major \Cyclus components are functioning as expected.

@@

Regression tests are tests which ensure that significant changes do not 
occur over the course of \Cyclus development. Such a change is called a 
\emph{regression} because the new version is almost always wrong.
Regression tests are implemented similarly to integration tests.
Nose is used to execute fully fledged \Cyclus simulations whose results
are then compared. In this category, however, the comparison is done against 
the output of the same input file when run with a previous version of \Cyclus, 
typically the last released version.
In some sense, regression tests are `dumb' in that they do 
not care about the contents of a simulation being correct, only whether or not 
it changed. They thus pick up glaring time-sensitive mistakes that may 
not appear elsewhere though they do not purport to make claims about the
quality of the physics modeled. The \Cycamore project is the primary location of
regression tests. This is because the archetypes here are sophisticated enough
to make interesting input files. Only a handful ($<10$) time steps are needed 
for regression testing.

@@

All of the public interface (the \gls{API}) must be documented as part of \Cyclus
\gls{QA} policy. This certifies that the intention for a code unit is communicated 
in way that is separate from its implementation. This policy is important 
because implementations are often wrong. Software implementations are unreliable 
resources about themselves. Truly trustworthy communication about what 
software should be doing comes from the author explicitly stating the 
goals of an \gls{API} in prose form. Future development of the code is then able to 
compare the implementation versus the documentation to discern whether or not they 
are in agreement. Documentation is thus a secondary higher-level information 
stream.  In \Cyclus, this information is aggregated together into static 
websites with the Doxygen \cite{van_heesch_doxygen:_2008} and Sphinx 
\cite{brandl_sphinx_2014} 
tools. These websites are placed online at \url{http://fuelcycle.org}.

@@

Version control is the history-preserving mechanism by which the provenance of 
a code is recored. \Cyclus uses a well-established tool called git 
\cite{software_freedom_conservancy_git_2014}
for this purpose. Git is currently the most popular utility in the recent wave
of \emph{distributed version control}. Briefly stated, this means that every 
user has a complete local copy of the \emph{repository} (the version control
term for the collection of all possible histories).
Though there and many praises and criticisms of git, one 
feature that the tool performs exceptionally well is the notion of code branches.
These are distinct pathways in the history that branch out of a mainline source
tree and then may be \emph{merged} back into the mainline. Multiple simultaneous
branches may exist at all point in time. Every change to the code is recorded
in the history, along with metadata such as the author, a timestamp, and an 
accompanying message. Thus 
it is possible for \Cyclus to accurately replay the entirety of who did what to the
code when.

\Cyclus uses a strategy known as \emph{git flow} 
\cite{kalliamvakou_code-centric_2014} 
to manage topical branches where individuals develop code, the mainline development
branch (simply known as develop), and the stable branch (called master).
In addition to these rules and delineation, \Cyclus also imposes rules on 
when it is acceptable from a quality control standpoint to merge code from 
the personal topical branches into develop. All software at this stage 
must have sufficient and passing tests and comprehensive documentation. Without
these core pieces proposed changes are not allowed into \Cyclus. A schematic of 
the development stages between reporting a bug and merging the fix is shown in 
Figure \ref{fig:gitprocess}.

\begin{figure}[htbp]
\begin{center}
\includegraphics[height=0.6\textheight]{./images/gitprocess}
\end{center}
\caption{A version-control enabled process for finding, reporting, fixing a bug 
can take place within an issue tracker alongside multiple stages of code 
testing and review. In the \Cyclus case, the final review stage includes unit, 
regression, and integration testing as well as manual code inspection.}
\label{fig:gitprocess}
\end{figure}

@@

In any multi-person software project, there is a tension between how individuals
wish to write code. Every person tends to have their own custom style. To remedy this
coding style guides are the software analogy to the natural language ones, 
such as the Chicago Manual of Style. \Cyclus strictly enforces the use of the 
Google C++ Style Guide \cite{weinberger_google_2008} for all software contributions.
This means that all developers of \Cyclus nominally write \Cyclus code in the same 
way.  This homogenization may be a hurdle to new developers but ultimately 
makes the it much more legible. Adoption of a single style guide also settles
timeless arguments over what style is `the best.' Having only one means that 
there is a common enemy for all developers to dislike equally.

@@

\emph{\acrfull{CI}} the idea that software should be tested and validated 
as it is being developed, rather than as a final stage in a longer development 
cycle.  Under \gls{CI} every pull request is tested independently immediately after 
it is proposed. Such tests should be run on all officially supported platforms. 
\Cyclus uses a \gls{CI} platform called Polyphemus \cite{scopatz_polyphemus_2014}.

@@

CI has become a necessary but not sufficient addition to the \Cyclus \gls{QA}
system. Its merit is that it keeps bad code out. However, \Cyclus will always
require human eyes and human hands to let good code in.
